\huge{Sexual selection improves population fitness: a systematic review and meta-analysis}

### 2019-03-07 Response_to_editorial_requests.Rmd

---
title: 'Sexual selection improves population fitness: a systematic review and meta-analysis'
output:
  pdf_document: default
subtitle: Response to Editorial Requests
bibliography: Manuscripts/Bibliography/Bibliography.bib
header-includes: 
- \usepackage{hyperref}\hypersetup{colorlinks=true, linkcolor=blue, filecolor=blue, urlcolor=blue} 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Dear Dr. Jones and colleagues,

We are delighted to provide a revised manuscript for publication in \textit{Nature Communications}. Below we provide a point-by-point response to the editorial requests as well as a response to comments by Prof. Jacek Radwan.

## Editorial Requests

\begin{enumerate}
  \item \textit{Please supply the main manuscript file in Microsoft Word or LaTeX format. Please supply a Supplementary Information file in PDF format.} \\ 
  The main manuscript is provided in LaTeX format (.tex) and the Supplementary Information is now in pdf format. However, we retain the HTML formatting for the Supplementary Information hosted on \hyperlink{https://justincally.github.io/SexualSelection/}{Github}.

  \item \textit{Please edit the title so that it is 15 words or fewer and does not include punctuation} \\
  <!-- Sexual selection improves population fitness based on a systematic review and meta-analysis -->

  \item \textit{The final paragraph of the Introduction should summarise the major results and conclusions of this manuscript, in the present tense.} \\
  <!-- We had previously included this in an earlier draft and cut it. Can add it back  -->

  \item \textit{Please shorten all subheadings in the Results section to fewer than 60 characters including spaces.} \\
  Sexual selection is associated with higher mean values for most fitness components \textbbf{to} Sexual selection improves average fitness for most traits. \\
  Sexual selection reduces phenotypic variance, for female traits in stressful environments \textbbf{to} Effects of sexual selection on phenotypic variance

  \item \textit{Please do not use italics or bold font to convey emphasis (in both the main text and the display items).} \\
  Is this the latin? (e.g. /Prima facie / Tabbles have bold, figures have bold...)

  \item \texit{Please avoid using speech marks around words or phrases. In most cases they are unnecessary.} \\
  Speech marks around the following words/phrases have now been removed: population fitness (l.21), purged (l.175), softer (l.199), harder (l.199), 'good genes' (l.243), attractive (l.244), unambiguous (l.359).

  \item \textit{Please make sure that mathematical terms throughout your manuscript and Supplementary Information (including in figures, figure axes, and legends) conform strictly to the guidelines} \\
  Not sure where we went wrong here?

  \item \textit{Wherever p-values are stated in the text and figure legends, please also state the name of the statistical test.} \\
  A couple of linear models in the publication bias section have p values without stating it is a linear regression.  Otherwise we dont have that many p values and it is described in the table as from a multilevel meta-analysis model.

  \item \textit{Please provide a full Methods section in the main manuscript file. Please note that there are no word limits to the Methods section. The Methods section must include subheadings of fewer than 60 characters including spaces. There is no word limit for the Methods section, so we recommend against splitting the methods description between the main text and Supplementary Information.}

  All headings are under 60 characters in length, we have some additional Supp Methods in the Supp info but it is more or an appendage to the Supp tables and search terms (which would not nicely format in the main text)

  \item \textit{Please rename the Methods section as "Methods".} \\

  Done

  \item \textit

\end{enumerate}

## Response to Prof. Jacek Radwan

### 2019-03-07 Manuscripts/Manuscript_MetaAnalysis.Rmd
Addition of editorial requests letter

---
title: '\huge{Sexual selection improves population fitness: a systematic review and meta-analysis}'
author: 
- Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- 
- '\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
subparagraph: yes
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
    \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
    \usepackage{hyperref}\hypersetup{colorlinks=false, linkcolor=blue, filecolor=blue, urlcolor=blue} 
    \usepackage{color,soul}
    \raggedbottom    
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict from first principles. Instead, one can empirically determine whether sexual selection increases or decreases population fitness by  manipulating the opportunity for sexual selection in replicate populations, allow evolution to proceed, and then compare the evolved average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 pertinent experimental evolution studies using meta-analysis. We find that sexual selection **on males** tends to elevate the mean and reduce the variance for many fitness traits, especially in females, and for populations evolving under stressful conditions. We conclude that the beneficial population-level consequences of sexual selection typically outweigh the harmful ones and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology. \vspace{5mm} \par\noindent \textbf{Keywords:} Adaptation, Gender load, Monogamy, Mutation load, Sexual conflict.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness, producing intralocus **sexual conflict [@RN187; @RN146; @RN237; @RN147; @Harano_2010; @Plesnar_2014]**. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males [@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care: \textit{inter}locus sexual conflict [@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness [@RN42; @RN25; @RN116; @RN95].

@@

Here, we synthesise the empirical literature on sexual selection and population fitness using formal meta-analysis. We focus exclusively on experimental evolution studies that manipulated the presence or strength of sexual selection **on males**, and then measured some fitness component, since experiments provide a particularly strong test of the hypothesis that sexual selection affects the average fitness of populations. Our principle aims were to measure the average net effect of sexual selection on the population-level mean and variance for various fitness components, to test whether this effect varies between stressful and benign environments, and identify key moderators of the effect size of sexual selection treatment. \phantomsection

@@

We retrieved 459 effect sizes from 65 studies. 92 effect sizes were collected from populations evolving under stressful conditions, while 337 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy ($n$ = 241); other effect sizes ($n$ = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 22 different fitness traits, with female reproductive success ($n$ = 102) and offspring viability ($n$ = 56) being the most commonly-measured traits. We classified **171 effect sizes as direct measures of population fitness, 141 as indirect, and the remaining 144 effect sizes as ambiguous** (see Methods, \autoref{forest_plot} and \href{https://justincally.github.io/SexualSelection/#inclusionexclusion_criteria}{Table S1}). Specifically, **we scored traits that are likely to correlate with population growth and persistence as direct** (*e.g.*, female reproductive success, offspring viability, and extinction rate), **those which are not necessarily correlated with population fitness but which do measure individual fitness as indirect** (*e.g.*, lifespan, male mating success and ejaculate quality/production); and **those for which the relationship to population fitness is unclear as ambiguous** (*e.g.*, body size, mating duration, male reproductive success, and early fecundity). \href{https://justincally.github.io/SexualSelection/#tables_of_sample_sizes}{Tables S4-S5} give a detailed description of our dataset.\phantomsection

@@

The grand mean across all types of effect sizes (**direct, indirect and ambiguous**) was positive (REML $\beta$ = 0.24, 95% CIs: 0.055 to 0.43, $p$ = 0.011; Bayesian $\beta$ = 0.25, 95% CIs: -0.0074 to 0.51, $BF_{>0}$ = 35), indicating that sexual selection **on males** typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, but varied across the three relationships to fitness \autoref{forest_plot}; \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S6}). **Sexual selection elevated fitness for traits that shared an ambiguous relationship to fitness (REML $\beta$ = 0.21, 95% CIs: 0.058 to 0.093; Bayesian $\beta$ = 0.20, 95% CIs: -0.0016 to	0.39, $BF_{>0}$ = 38, $n$ = 144) and an indirect relationship to fitness (REML $\beta$ = 0.24, 95% CIs: 0.13 to 0.36; Bayesian $\beta$ = 0.24, 95% CIs: 0.033 to 0.43, $BF_{>0}$ = 59, $n$ = 141). Additionally, sexual selection elevated fitness components directly related to fitness, albeit at a lower magnitude (REML $\beta$ = 0.13, 95% CIs: 0.019 to 0.24; Bayesian $\beta$ = 0.13, 95% CIs: -0.079 to 0.31, $BF_{>0}$ = 11, $n$ = 174). A large forest plot with predicted effect sizes for each fitness component is presented in Figure S1 and further detailed with model predictions and individual meta-analyses in** \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S8 and S9}. Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.42, 95% CIs: -0.64 to -0.20; Bayesian $\beta$ = -0.43, 95% CIs: -0.70 to -0.15, $BF_{>0}$ = 0.0026; $n$ = 35) and body condition (REML $\beta$ = -1.2, 95% CIs: -1.8 to -0.63; Bayesian $\beta$ = -1.2, 95% CIs: -1.9 to -0.63, $BF_{>0}$ < 0.0001, $n$ = 1).  \phantomsection

@@

We found that the sex of the individuals measured (male, female, or a mixture), and the conditions under which the population evolved (stressful or benign) interacted to affect the relationship between sexual selection and fitness (\hyperlink{page.18}{Table 1, 2}). Sexual selection **on males** significantly improved female fitness, and the beneficial effect of sexual selection was significantly stronger for females from populations evolving under stressful conditions (*e.g.* a food source to which they were not well-adapted) than under benign conditions (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Sexual selection had a positive but non-significant effect on male fitness, and in contrast to females, fitness benefits were significantly weaker in stressful than benign environments (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Consistent with the different consequences of sexual selection for female and male fitness, the mean effect size in mixed-sex samples was non-significantly positive, and there was no significant difference between benign versus stressful conditions (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Overall, our results indicate that the positive effect of sexual selection on fitness is greater for females than males, and the difference between the sexes is magnified in stressful environments. **When we used an alternative measure of effect size (log response ratio or lnRR), the results aligned with those of Hedges' g in that sexual selection elevates population fitness, with its effect magnified for females evolving in stressful environments (Figure S3 and Table S16).**

@@

The effect size estimates we recovered were highly heterogeneous **($I^2$ = 95.2%, 95% CIs: 94.4% to 95.9%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208; @Senior_I2_2016]. Heterogeneity stemmed mostly from between-study differences ($I^2_{study}$ = 36%, 95% CIs: 26.5% to 45.4%) rather than differences between fitness components and taxon ($I^2_{fitness\:components}$ = 0.4%, 95% CIs: 0.2% to 0.9%; $I^2_{taxon}$ = 1.4%, 95% CIs: 0.2% to 3.5%).** Variation among taxa is explored further in the ESM (\href{https://justincally.github.io/SexualSelection/#restricted_analysis_to_just_the_higher-quality_data}{Figure S2, Table S11}).  \phantomsection

@@

By applying meta-analysis to log coefficient of variation ratios [*lnCVR*; @RN177], we found evidence that sexual selection reduces phenotypic variation under certain conditions (\autoref{env_sex}b). Specifically, phenotypic variance was significantly reduced by sexual selection for fitness components measured in females under stressful conditions (*lnCVR* = -0.78, 95% CIs: -1.23 to -0.34, $n$ = 27). By contrast, we found no significant effect of sexual selection on phenotypic variance in males, or for either sex under benign conditions (\autoref{env_sex}b; \href{https://justincally.github.io/SexualSelection/#obtaining_lncvr_and_meta-analysis_models}{Tables S13-S16}). However, similar to the results in females, there was a non-significant trend for a reduction in phenotypic variance in mixed-sex samples measured under stressful conditions (*lnCVR* = -0.76, 95% CIs: -1.22 to -0.31; \autoref{env_sex}b). **A meta-analysis using the log variability ratio (lnVR), which does not account for the mean-variance relationship present in the dataset (Figure S4), suggested sexual selection reduces variance for mixed-sex samples in stressful conditions, but not females, results of this meta-analysis can be found in the Supplementary Material (Figure S5 and Table S21)**.

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 98.9%, 95% CIs: 98.7% to 99.1%). **Heterogeneity in the dataset was due to variability between studies fitness components ($I^2_{fitness\:components}$ = 12.7%, 95% CIs: 4.5% to 23.1%) and taxon ($I^2_{taxon}$ = 7.7%, 95% CIs: 1.3% to 18.4%), as well as large amounts of residual heterogeneity (78.6 %, 95 % CIs: 65.9% to 89%). Using the REML approach heterogeneity associated with the Study ID random effect was estimated at zero ($I^2_{study}$ = 0 %), however the Bayesian approach suggests heterogeneity between studies may be zero or small ($sd_{study}$ = 0.07, 95 % CIs: 0 to 0.19).**  \phantomsection

@@

The funnel plot of effect sizes was asymmetrical, suggesting that some publication bias might be present (\autoref{publication_bias}a; Egger’s test: *z* = 5.9, $p$ < 0.0001). Specifically, there was a moderate excess of low-powered studies in which sexual selection had a more positive effect on the fitness component than average, implying that low-powered studies are more likely to be published if they report statistically significant fitness benefits of sexual selection **(though funnel plots are not a decisive evidence of publication bias[@RN215])**. There was no significant relationship between effect size and journal impact factor (\autoref{publication_bias}b; $t_{437}$ = 1.2, $p$ = 0.23) or year of publication (\autoref{publication_bias}c; $t_{437}$ = -1.2, $p$ = 0.24); thus, we found no evidence that effect size dictates the likelihood of publication in high-profile journals, or that effect sizes have diminished as the field has matured [@RN216; @RN217].

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection **on males** was experimentally removed or weakened. **Sexual selection had beneficial effects for traits that were ambiguously or indirectly related to fitness, but had smaller, variable effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced) and the proportion of viable offspring**. Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. This result is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example, due to immunosuppressive effects of sex hormones [@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males, and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, **female and mixed-sex samples showed reduced phenotypic variance** when sexual selection was applied under stressful as opposed to benign conditions. These results suggest that sexual selection may contribute to population persistence under stressful conditions, such as fluctuating environmental change [@RN173] or spatial variability [@RN189], particularly since female reproductive output is often a limiting factor in population growth [@RN188].

The results of the meta-analysis support predictions that sexual selection **on males** can improve population fitness and accelerate adaptation [@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. Specifically, our results support arguments that positive genetic correlations between male mating success and female non-sexual traits are common, allowing females to benefit from a genome that has been 'purged' of deleterious alleles **through sexual selection on males** [@RN94; @RN198; @RN95]. What is less clear is why sexual selection had a larger effect on female trait means and variances as opposed to males (the sex in which sexual selection is typically strongest). **Below we discuss possible explanations for this result, in light of the core principle that the extent of adaptation depends on additive genetic variance, the structure of genetic covariance, and the strength of selection [@RN317; @RN318]. However, these explanations are speculative and _post hoc_, as the result is arguably opposite to what one would predict, (*i.e.* that manipulating sexual selection on males should affect the phenotype of males more than females)**.

@@

#Materials and methods
\phantomsection

@@

From each paper, we first attempted to extract the arithmetic means, standard deviations and sample sizes of each of the different treatment groups, which facilitate calculation of effect size (see below). Typically, there were two or three treatments, which varied in the strength of sexual selection **on males through manipulations to the adult sex ratio, in these cases we considered treatments with the greater male to female ratio to be the high sexual selection treatment group**. For some papers, summary statistics were not written down, but were presented in a figure such as a bar chart: in these cases, we extracted the data using *WebPlotDigitizer* v.3.12 [@RN195]. If the treatment means were not reported (and the raw data were unavailable), we instead calculated effect size from test statistics comparing treatment means (*e.g.*, $F$, $t$, $z$ or $\chi^2$ values), which we used to estimate effect size using several formulae (see below).

@@

Firstly, we obtained a weighted mean effect size (Hedges’ *g*) for the entire dataset, using both Bayesian and restricted maximum likelihood (REML) approaches for completeness. The weighted mean was obtained by fitting a model with no moderator variables (*i.e.* fixed effects), but fitness component (*e.g.*, body size, female reproductive success), study ID, and taxon as random/group-level effects. That is, we separately model correlations between different effect sizes sourced from the same study, taxon, or pertaining to the same fitness component, and account for these interdependencies when estimating the overall effect. Given the small number of phylogenetically diverse species, we did not utilise phylogenetic corrections within the models. **In our meta-analyses we report Bayes Factors ($BF$), giving the likelihood ratio that the focal effect size differs from zero $BF_{>0}$.**

>>>>

---
title: '\huge{Sexual selection improves population fitness: a systematic review and meta-analysis}'
author: 
- Justin G. Cally\textsuperscript{§*}, Devi Stuart-Fox\textsuperscript{§} and Luke Holman\textsuperscript{§}
- 
- '\textsuperscript{*}\textit{justin.g.cally@gmail.com}'
output:
  pdf_document: default
  html_document: default
  word_document:
      pandoc_args: ["--smart"]
bibliography: Bibliography/Bibliography.bib
subparagraph: yes
csl: Bibliography/NComms.csl
header-includes: \usepackage{microtype} \usepackage[all]{hypcap} \usepackage{amsmath}
  \usepackage{booktabs} \usepackage{caption} \usepackage[labelfont=bf]{caption} \usepackage{titling}
  \pretitle{\begin{flushleft}\LARGE} \posttitle{\par\end{flushleft}\vskip 0.5em} \preauthor{\begin{flushleft}\large
  \lineskip 0.5em} \postauthor{\par\end{flushleft}} \predate{\begin{flushleft}\large}
  \postdate{\par\end{flushleft}} \usepackage{fancyhdr} \pagestyle{fancy} \fancyhead[LO,LE]{\textsl{\leftmark}}
  \rhead[]{Sexual selection improves fitness} \usepackage{titlefoot}
  \usepackage{amssymb} \usepackage{rotating} \usepackage{lineno} \linenumbers \usepackage{setspace}\onehalfspacing
    \usepackage{titlesec} \titleformat{\section}{\normalfont\scshape\LARGE}{\thesection}{1em}{} \titleformat{\subsection}{\normalfont\Large}{\thesubsection}{1em}{} \titleformat{\subsubsection}{\normalfont\large}{\thesubsubsection}{1em}{} \usepackage{abstract} \renewcommand{\abstractnamefont}{\normalfont\Large} 
    \usepackage{hyperref}\hypersetup{colorlinks=false, linkcolor=blue, filecolor=blue, urlcolor=blue} 
    \usepackage{color,soul}
    \raggedbottom    
abstract: Sexual selection has manifold ecological and evolutionary consequences, making its net effect on population fitness difficult to predict from first principles. Instead, one can empirically determine whether sexual selection increases or decreases population fitness by  manipulating the opportunity for sexual selection in replicate populations, allow evolution to proceed, and then compare the evolved average fitness between the treatment groups. Here, we synthesise 459 effect sizes from 65 pertinent experimental evolution studies using meta-analysis. We find that sexual selection on males tends to elevate the mean and reduce the variance for many fitness traits, especially in females, and for populations evolving under stressful conditions. We conclude that the beneficial population-level consequences of sexual selection typically outweigh the harmful ones and that the effects of sexual selection on total selection, genetic variance, and/or evolvability can differ markedly between sexes and environments. We discuss the implications of these results for conservation and evolutionary biology. \vspace{5mm} \par\noindent \textbf{Keywords:} Adaptation, Gender load, Monogamy, Mutation load, Sexual conflict.
---
\maketitle\unmarkedfntext{\textsuperscript{§}School of BioSciences, The University of Melbourne, Victoria, Australia.}

@@

Conversely, sexual selection can decrease population fitness if male sexually-selected traits are negatively genetically correlated with female fitness, producing intralocus sexual conflict [@RN187; @RN146; @RN237; @RN147; @Harano_2010; @Plesnar_2014]. Additionally, sexual selection frequently favours phenotypes that reduce population fitness but benefit the individuals expressing them, such as harassment or infanticide by mate-seeking males [@RN6; @RN267; @RN37], as well as investment in costly sexual signals and weaponry at the expense of parental care: \textit{inter}locus sexual conflict [@RN314; @RN276; @RN315]. Given these conflicting theoretical expectations and empirical results, it remains unclear whether sexual selection tends to have a net benefit or cost to population fitness [@RN42; @RN25; @RN116; @RN95].

@@

Here, we synthesise the empirical literature on sexual selection and population fitness using formal meta-analysis. We focus exclusively on experimental evolution studies that manipulated the presence or strength of sexual selection on males, and then measured some fitness component, since experiments provide a particularly strong test of the hypothesis that sexual selection affects the average fitness of populations. Our principle aims were to measure the average net effect of sexual selection on the population-level mean and variance for various fitness components, to test whether this effect varies between stressful and benign environments, and identify key moderators of the effect size of sexual selection treatment. \phantomsection

@@

We retrieved 459 effect sizes from 65 studies. 92 effect sizes were collected from populations evolving under stressful conditions, while 337 were measured for those evolving in benign conditions. 189 of the effect sizes came from measurements made on males, 219 on females, and the remaining 51 from measurements of a mixed-sex sample of individuals. Most effect sizes in our dataset came from studies that manipulated sexual selection by completely removing it in one treatment via enforced random monogamy ($n$ = 241); other effect sizes ($n$ = 218) derived from alternative manipulations, such as changing the adult sex ratio. In total, we obtained effect sizes for 22 different fitness traits, with female reproductive success ($n$ = 102) and offspring viability ($n$ = 56) being the most commonly-measured traits. We classified 171 effect sizes as direct measures of population fitness, 141 as indirect, and the remaining 144 effect sizes as ambiguous (see Methods, \autoref{forest_plot} and \href{https://justincally.github.io/SexualSelection/#inclusionexclusion_criteria}{Table S1}). Specifically, we scored traits that are likely to correlate with population growth and persistence as direct (*e.g.*, female reproductive success, offspring viability, and extinction rate), those which are not necessarily correlated with population fitness but which do measure individual fitness as indirect (*e.g.*, lifespan, male mating success and ejaculate quality/production); and those for which the relationship to population fitness is unclear as ambiguous (*e.g.*, body size, mating duration, male reproductive success, and early fecundity). \href{https://justincally.github.io/SexualSelection/#tables_of_sample_sizes}{Tables S4-S5} give a detailed description of our dataset.\phantomsection

@@

The grand mean across all types of effect sizes (direct, indirect and ambiguous) was positive (REML $\beta$ = 0.24, 95% CIs: 0.055 to 0.43, $p$ = 0.011; Bayesian $\beta$ = 0.25, 95% CIs: -0.0074 to 0.51, $BF_{>0}$ = 35), indicating that sexual selection on males typically had a net positive effect on the majority of populations and fitness components so far studied. Moreover, the effect sizes associated with the manipulation of sexual selection varied between different fitness traits. Sexual selection had a beneficial effect on most fitness traits, but varied across the three relationships to fitness \autoref{forest_plot}; \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S6}). Sexual selection elevated fitness for traits that shared an ambiguous relationship to fitness (REML $\beta$ = 0.21, 95% CIs: 0.058 to 0.093; Bayesian $\beta$ = 0.20, 95% CIs: -0.0016 to	0.39, $BF_{>0}$ = 38, $n$ = 144) and an indirect relationship to fitness (REML $\beta$ = 0.24, 95% CIs: 0.13 to 0.36; Bayesian $\beta$ = 0.24, 95% CIs: 0.033 to 0.43, $BF_{>0}$ = 59, $n$ = 141). Additionally, sexual selection elevated fitness components directly related to fitness, albeit at a lower magnitude (REML $\beta$ = 0.13, 95% CIs: 0.019 to 0.24; Bayesian $\beta$ = 0.13, 95% CIs: -0.079 to 0.31, $BF_{>0}$ = 11, $n$ = 174). A large forest plot with predicted effect sizes for each fitness component is presented in Figure S1 and further detailed with model predictions and individual meta-analyses in \href{https://justincally.github.io/SexualSelection/#effect_of_sexual_selection_for_different_fitness_components}{Table S8 and S9}. Sexual selection significantly reduced two fitness components, namely immunity (REML $\beta$ = -0.42, 95% CIs: -0.64 to -0.20; Bayesian $\beta$ = -0.43, 95% CIs: -0.70 to -0.15, $BF_{>0}$ = 0.0026; $n$ = 35) and body condition (REML $\beta$ = -1.2, 95% CIs: -1.8 to -0.63; Bayesian $\beta$ = -1.2, 95% CIs: -1.9 to -0.63, $BF_{>0}$ < 0.0001, $n$ = 1).  \phantomsection

@@

We found that the sex of the individuals measured (male, female, or a mixture), and the conditions under which the population evolved (stressful or benign) interacted to affect the relationship between sexual selection and fitness (\hyperlink{page.18}{Table 1, 2}). Sexual selection on males significantly improved female fitness, and the beneficial effect of sexual selection was significantly stronger for females from populations evolving under stressful conditions (*e.g.* a food source to which they were not well-adapted) than under benign conditions (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Sexual selection had a positive but non-significant effect on male fitness, and in contrast to females, fitness benefits were significantly weaker in stressful than benign environments (\autoref{env_sex}a, \hyperlink{page.18}{Table 2}). Consistent with the different consequences of sexual selection for female and male fitness, the mean effect size in mixed-sex samples was non-significantly positive, and there was no significant difference between benign versus stressful conditions (\autoref{env_sex}a, \hyperlink{page.13}{Table 2}). Overall, our results indicate that the positive effect of sexual selection on fitness is greater for females than males, and the difference between the sexes is magnified in stressful environments. When we used an alternative measure of effect size (log response ratio or lnRR), the results aligned with those of Hedges' g in that sexual selection elevates population fitness, with its effect magnified for females evolving in stressful environments (Figure S3 and Table S16).

@@

The effect size estimates we recovered were highly heterogeneous ($I^2$ = 95.2%, 95% CIs: 94.4% to 95.9%), reflecting the large differences in experimental procedures, study species, and fitness components included in our meta-analysis [@RN208; @Senior_I2_2016]. Heterogeneity stemmed mostly from between-study differences ($I^2_{study}$ = 36%, 95% CIs: 26.5% to 45.4%) rather than differences between fitness components and taxon ($I^2_{fitness\:components}$ = 0.4%, 95% CIs: 0.2% to 0.9%; $I^2_{taxon}$ = 1.4%, 95% CIs: 0.2% to 3.5%). Variation among taxa is explored further in the ESM (\href{https://justincally.github.io/SexualSelection/#restricted_analysis_to_just_the_higher-quality_data}{Figure S2, Table S11}).  \phantomsection

@@

By applying meta-analysis to log coefficient of variation ratios [*lnCVR*; @RN177], we found evidence that sexual selection reduces phenotypic variation under certain conditions (\autoref{env_sex}b). Specifically, phenotypic variance was significantly reduced by sexual selection for fitness components measured in females under stressful conditions (*lnCVR* = -0.78, 95% CIs: -1.23 to -0.34, $n$ = 27). By contrast, we found no significant effect of sexual selection on phenotypic variance in males, or for either sex under benign conditions (\autoref{env_sex}b; \href{https://justincally.github.io/SexualSelection/#obtaining_lncvr_and_meta-analysis_models}{Tables S13-S16}). However, similar to the results in females, there was a non-significant trend for a reduction in phenotypic variance in mixed-sex samples measured under stressful conditions (*lnCVR* = -0.76, 95% CIs: -1.22 to -0.31; \autoref{env_sex}b). A meta-analysis using the log variability ratio (lnVR), which does not account for the mean-variance relationship present in the dataset (Figure S4), suggested sexual selection reduces variance for mixed-sex samples in stressful conditions, but not females, results of this meta-analysis can be found in the Supplementary Material (Figure S5 and Table S21).

As in the meta-analysis of trait means, there was high heterogeneity in the estimates of *lnCVR* ($I^2$ = 98.9%, 95% CIs: 98.7% to 99.1%). Heterogeneity in the dataset was due to variability between studies fitness components ($I^2_{fitness\:components}$ = 12.7%, 95% CIs: 4.5% to 23.1%) and taxon ($I^2_{taxon}$ = 7.7%, 95% CIs: 1.3% to 18.4%), as well as large amounts of residual heterogeneity (78.6 %, 95 % CIs: 65.9% to 89%). Using the REML approach heterogeneity associated with the Study ID random effect was estimated at zero ($I^2_{study}$ = 0 %), however the Bayesian approach suggests heterogeneity between studies may be zero or small ($sd_{study}$ = 0.07, 95 % CIs: 0 to 0.19).  \phantomsection

@@

The funnel plot of effect sizes was asymmetrical, suggesting that some publication bias might be present (\autoref{publication_bias}a; Egger’s test: *z* = 5.9, $p$ < 0.0001). Specifically, there was a moderate excess of low-powered studies in which sexual selection had a more positive effect on the fitness component than average, implying that low-powered studies are more likely to be published if they report statistically significant fitness benefits of sexual selection (though funnel plots are not a decisive evidence of publication bias[@RN215]). There was no significant relationship between effect size and journal impact factor (\autoref{publication_bias}b; $t_{437}$ = 1.2, $p$ = 0.23) or year of publication (\autoref{publication_bias}c; $t_{437}$ = -1.2, $p$ = 0.24); thus, we found no evidence that effect size dictates the likelihood of publication in high-profile journals, or that effect sizes have diminished as the field has matured [@RN216; @RN217].

@@

Our meta-analysis revealed that populations evolving under sexual selection often have higher values for multiple fitness traits, relative to populations where sexual selection on males was experimentally removed or weakened. Sexual selection had beneficial effects for traits that were ambiguously or indirectly related to fitness, but had smaller, variable effects on direct measures of fitness such as extinction rate, reproductive success (defined as the number of offspring produced) and the proportion of viable offspring. Fitness traits related to immunocompetence were an exception: sexual selection typically resulted in weaker immunity. This result is interesting in light of the hypothesised trade-off between sexually-selected phenotypes and immunity, for example, due to immunosuppressive effects of sex hormones [@RN323; @RN324]. Furthermore, the overall benefit of sexual selection was greater for females than males, and this sex difference was magnified in stressful environments. Consistent with stronger selection on female fitness under stress, female and mixed-sex samples showed reduced phenotypic variance when sexual selection was applied under stressful as opposed to benign conditions. These results suggest that sexual selection may contribute to population persistence under stressful conditions, such as fluctuating environmental change [@RN173] or spatial variability [@RN189], particularly since female reproductive output is often a limiting factor in population growth [@RN188].

The results of the meta-analysis support predictions that sexual selection on males can improve population fitness and accelerate adaptation [@RN94; @RN173; @RN54; @RN141; @RN230; @RN198; @RN58]. Specifically, our results support arguments that positive genetic correlations between male mating success and female non-sexual traits are common, allowing females to benefit from a genome that has been 'purged' of deleterious alleles through sexual selection on males [@RN94; @RN198; @RN95]. What is less clear is why sexual selection had a larger effect on female trait means and variances as opposed to males (the sex in which sexual selection is typically strongest). Below we discuss possible explanations for this result, in light of the core principle that the extent of adaptation depends on additive genetic variance, the structure of genetic covariance, and the strength of selection [@RN317; @RN318]. However, these explanations are speculative and _post hoc_, as the result is arguably opposite to what one would predict, (*i.e.* that manipulating sexual selection on males should affect the phenotype of males more than females).

@@

#Methods
\phantomsection

@@

From each paper, we first attempted to extract the arithmetic means, standard deviations and sample sizes of each of the different treatment groups, which facilitate calculation of effect size (see below). Typically, there were two or three treatments, which varied in the strength of sexual selection on males through manipulations to the adult sex ratio, in these cases we considered treatments with the greater male to female ratio to be the high sexual selection treatment group. For some papers, summary statistics were not written down, but were presented in a figure such as a bar chart: in these cases, we extracted the data using *WebPlotDigitizer* v.3.12 [@RN195]. If the treatment means were not reported (and the raw data were unavailable), we instead calculated effect size from test statistics comparing treatment means (*e.g.*, $F$, $t$, $z$ or $\chi^2$ values), which we used to estimate effect size using several formulae (see below).

@@

Firstly, we obtained a weighted mean effect size (Hedges’ *g*) for the entire dataset, using both Bayesian and restricted maximum likelihood (REML) approaches for completeness. The weighted mean was obtained by fitting a model with no moderator variables (*i.e.* fixed effects), but fitness component (*e.g.*, body size, female reproductive success), study ID, and taxon as random/group-level effects. That is, we separately model correlations between different effect sizes sourced from the same study, taxon, or pertaining to the same fitness component, and account for these interdependencies when estimating the overall effect. Given the small number of phylogenetically diverse species, we did not utilise phylogenetic corrections within the models. In our meta-analyses we report Bayes Factors ($BF$), giving the likelihood ratio that the focal effect size differs from zero $BF_{>0}$.
